Marlowe: implementing and analysing financial contracts on blockchain

### 2019-12-12 papers/marlowe2019/paper.tex
15pp

Marlowe\footnote{The complete Marlowe project, including semantics and implementation, is available at \url{https://github.com/input-output-hk/marlowe}.} is a domain-specific language (DSL) for implementing financial contracts on blockchain: our initial target is Cardano, but it could be implemented on many distributed ledgers (DLT platforms), including Ethereum. Marlowe is embedded in Haskell, allowing users selectively to use aspects of Haskell -- typically definitions of constants and simple functions -- to express contracts more readably and succinctly. Section \ref{sec:overview} gives an overview of the language, and the changes made to it since it was originally introduced in ~\cite{isola-marlowe}.

@@

Contracts are built by putting together a small number of constructs that in combination can be used to describe many different kinds of financial contract.

Contracts in Marlowe run on a blockchain, but need to interact with the off-chain world. The parties to the contract, also called the participants, can engage in various actions: they can be asked to deposit money, or to make a choice between various alternatives. In some cases, any party will be able to trigger the contract just to notify it that some condition has become true (for example, a timeout has occurred).

@@

Cardano aims to support smart contracs during its Shelley release in 2020.
Cardano smart contract platform is called \emph{Plutus}, and
it uses Haskell programming language to generate a form of $System F_{\omega}$,
called \emph{Plutus Core}, by extending GHC using its plugin
support~\cite[Section 13.3]{ghcusersguide}.

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=4in]{figures/Marlowe3-Figure-1-crop.pdf}
    \caption{Cardano EUTx0 Model}
    \label{fig:eutxo}
\end{figure}

@@

We then compute expected transaction outcomes for given inputs, such as:
contract continuation, new state, and payments.

\begin{verbatim}
expectedTxOutputs = computeTransaction txInput marloweState marloweContract
\end{verbatim}

\noindent
where

\begin{verbatim}
computeTransaction :: TransactionInput -> State -> Contract -> TransactionOutput

data TransactionInput = TransactionInput
    { txInterval :: SlotInterval
    , txInputs   :: [Input] }

data TransactionOutput =
    TransactionOutput
        { txOutWarnings :: [TransactionWarning]
        , txOutPayments :: [Payment]
        , txOutState    :: State
        , txOutContract :: Contract }
    | Error TransactionError
\end{verbatim}

\noindent
Given a list of \texttt{Input}'s from \emph{Redeemer}, the interpreter reduces a contract until it becomes quiescent: either it evaluates to \texttt{Close},
or it expects a user input in a \texttt{When} construct.
All \texttt{Pay, If, Let, Close} constructs are evaluated immediately.

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=4in]{figures/Marlowe3-Figure-2-crop.pdf}
    \caption{Zero Coupon Bond Contract}
    \label{fig:zero-coupon-bond}
\end{figure}

@@

\subsection{Further Work}

Cardano extends its ledger rules to support \emph{forging} of custom currencies and tokens,
effectively embedding support of Ethereum ERC-20 and ERC-721 contracts. 
Simple token creation gives interesting possibilities of representing Marlowe contract parties by tokens.
This tokenization of contract participants abstracts away concrete public keys into contract \emph{roles}.
In turn, those roles could be traded independently of a contract.

We are working on adding \emph{multicurrency} or \emph{roles} support to Marlowe.

@@

Another property of Marlowe is that any given finite \texttt{Contract} has an implicit finite bound on the maximum number of \texttt{Transaction}s that it accepts. This is a convenient property because:

\begin{itemize}
    \item It reduces the danger of Denial of Service (DoS) attacks: because the number of valid inputs is limited, an attacker participant cannot arbitrarily block the contract by issuing an unbounded amount of useless \texttt{Transaction}s.
    \item The number of transactions bounds the length of traces that symbolic execution (see Section~\ref{sec:static}) needs to explore (more on this below).%SJT: is this really below? Or above?
\end{itemize}

We state the property as follows:

@@

\begin{wrapfigure}[9]{l}{0.25\textwidth}
    \vspace*{-0.28in}
    \includegraphics[width=0.25\textwidth]{hole_options.png}
\end{wrapfigure}

@@

\begin{wrapfigure}[6]{l}{0.25\textwidth}
    \vspace*{-0.00in}
    \includegraphics[scale=0.2]{hole_options_2.png}
\end{wrapfigure}
A complete contract can be written in this guided way with the user needing only to fill in strings and numbers by hand. This approach to writing holes in your code and ``asking'' the compiler what you could put in there is easy to implement in a DSL because there are very limited options, however is is also becoming popular with more complex languages such as Haskell and Idris.

@@

\noindent
The user can then choose to add this action to a transaction being prepared. Once the action is added other inputs become possible; these are displayed in the input composer, and again they can be added to the transaction being composed. In this way, multiple actions can be added to a transaction before it is applied.

\begin{wrapfigure}[6]{l}{0.5\textwidth}
    \vspace*{-0.2in}
    \includegraphics[scale=0.15]{tx_composer.png}
\end{wrapfigure}

\noindent
A user can then apply this transaction and in the example above this would result in the state pane showing a single payment and in addition the contract in the Marlowe editor will have been reduced to \lstinline{Close}.

@@

An earlier paper~\cite{cryptoeprint:2016:1156} reviews smart contracts on blockchain. Here we look at a number of recent systems that bear direct comparison with Marlowe.

The Ethereum system~\cite{wood2014ethereum} provides an exemplar Turing-complete language for  the EVM virtual machine, and a higher-level programming language, Solidity, that compiles into EVM code. Both these languages are general-purpose, and so have more akin with Plutus~\cite{PlutusPlatform} rather than Marlowe.

On the other hand,  Nxt~\cite{Nxt}, is special-purpose in providing a ``fat'' high-level
API, containing built-in transaction types and transactions that support some 250 primitive operations; these can be ``scripted'' in a client (only) using a binding to the API, which is available, for instance, in JavaScript. In providing such specificity this bears comparison with our implementation of contracts from the ACTUS standard~\cite{actus}.

Our work is inspired by the original work of Peyton Jones and others~\cite{PeytonJones:2000} to describe financial contracts using a DSL embedded in Haskell. The Findel project~\cite{findel} examines financial contracts on the Ethereum platform, and is also based on~\cite{PeytonJones:2000}. The authors note that payments need to be bounded; this is made concrete in our
account by our notion of commitments. They take no account of commitments or timeouts as our approach does, and so are unable to guarantee some of the properties -- such as a finite lifetime -- built into Marlowe by design.

BitML~\cite{BitML} is a DSL for specifying Marlowe-like contracts that regulate transfers on the Bitcoin blockchain. BitML is implemented via a compiler that translates contracts into standard Bitcoin transactions. Participants execute a contract by appending these transactions on the Bitcoin blockchain, according to their strategies, which also involve the exchange of bitstrings that guarantee to a very high probability the correctness of contract execution. Marlowe is directly implemented by an interpreter which could also be implemented on a covenant-based~\cite{covenants} extension of the Bitcoin blockchain.

>>>>

Marlowe\footnote{see \url{https://github.com/input-output-hk/marlowe} for complete project} is a domain-specific language (DSL) for implementing financial contracts on blockchain: our initial target is Cardano, but it could be implemented on many distributed ledgers (DLT platforms), including Ethereum. Marlowe is embedded in Haskell, allowing users selectively to use aspects of Haskell -- typically definitions of constants and simple functions -- to express contracts more readably and succinctly. Section \ref{sec:overview} gives an overview of the language, and the changes made to it since it was originally introduced in ~\cite{isola-marlowe}.

@@

Contracts are built by putting together a small number of constructs that in combination can  describe many different financial contracts.

%Contracts in Marlowe run on a blockchain, but need to interact with the off-chain world. 
The parties to the contract, also called the participants, can engage in various actions: they can be asked to deposit money, or to make a choice between various alternatives. In some cases, any party will be able to trigger the contract just to notify it that some condition has become true (e.g., a timeout has occurred).

@@

Cardano aims to support smart contracts during its Shelley release in 2020.
Cardano smart contract platform is called \emph{Plutus}, and
it uses Haskell programming language to generate a form of $System F_{\omega}$,
called \emph{Plutus Core}, by extending GHC using its plugin
support~\cite[Section 13.3]{ghcusersguide}.

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=4in]{figures/Marlowe3-Figure-1-crop.pdf}
    %\caption{Cardano EUTx0 Model}
    %\label{fig:eutxo}
\end{figure}

@@

We then apply  \texttt{computeTransaction}  to the contract inputs, the contract continuation, and new state to compute the expected transaction outcomes:
\begin{verbatim}
computeTransaction :: 
  TransactionInput -> State -> Contract -> TransactionOutput
\end{verbatim}
%\begin{verbatim}
%expectedTxOutputs = computeTransaction txInput marloweState marloweContract
%\end{verbatim}
%
%\noindent
%where
%
%\begin{verbatim}
%computeTransaction :: TransactionInput -> State -> Contract -> TransactionOutput
%
%data TransactionInput = TransactionInput
%    { txInterval :: SlotInterval
%    , txInputs   :: [Input] }
%
%data TransactionOutput =
%    TransactionOutput
%        { txOutWarnings :: [TransactionWarning]
%        , txOutPayments :: [Payment]
%        , txOutState    :: State
%        , txOutContract :: Contract }
%    | Error TransactionError
%\end{verbatim}
\noindent
where a \texttt{TransactionInput} consists of the current slot interval, together with other ontract inputs, and the outputs combine any payments and warnings with the resulting output state and contract.

Given a list of \texttt{Input}'s from \emph{Redeemer}, the interpreter reduces a contract until it becomes quiescent: either it evaluates to \texttt{Close},
or it expects a user input in a \texttt{When} construct.
All \texttt{Pay, If, Let, Close} constructs are evaluated immediately.

@@

\begin{figure}[!h]
    \centering
    \includegraphics[width=4in]{figures/Marlowe3-Figure-2-crop.pdf}
    %\caption{Zero Coupon Bond Contract}
    %\label{fig:zero-coupon-bond}
\end{figure}

@@

\paragraph{Future Work}

Cardano extends its ledger rules to support \emph{forging} of custom currencies and tokens.%effectively embedding support of Ethereum ERC-20 and ERC-721 contracts. 
Simple token creation gives interesting possibilities of representing Marlowe contract parties by tokens.
This tokenization of contract participants abstracts away concrete public keys into contract \emph{roles}.
In turn, those roles could be traded independently of a contract. 
We are working on adding \emph{multicurrency} or \emph{roles} support to Marlowe.

@@

Another property of Marlowe is that any given \texttt{Contract} has an implicit finite bound on the maximum number of \texttt{Transaction}s that it accepts. 
This is a convenient property for two reasons.

First, it reduces the danger of Denial of Service (DoS) attacks, because the number of valid inputs is limited, an attacker participant cannot arbitrarily block the contract by issuing an unbounded amount of useless \texttt{Transaction}s. Secondly, the number of transactions bounds the length of traces that symbolic execution (see Section~\ref{sec:static}) needs to explore.% (more on this below).%SJT: is this really below? Or above?
We state the property as follows:

@@

\begin{wrapfigure}[9]{l}{0.25\textwidth}
    %\vspace*{-0.28in}
    \includegraphics[width=0.25\textwidth]{hole_options.png}
\end{wrapfigure}

@@

\begin{wrapfigure}[6]{l}{0.25\textwidth}
    \vspace*{-0.25in}
    \includegraphics[scale=0.2]{hole_options_2.png}
\end{wrapfigure}
A complete contract can be written in this guided way with the user needing only to fill in strings and numbers by hand. This approach to writing holes in your code and ``asking'' the compiler what you could put in there is easy to implement in a DSL because there are very limited options, however is is also becoming popular with more complex languages such as Haskell and Idris.

@@

\noindent
The user can then choose to add this action to a transaction being prepared. Once the action is added other inputs become possible; these are displayed in the input composer, and again they can be added to the transaction being composed. In this way, multiple %actions can be added to a transaction before it is applied.

\begin{wrapfigure}[5]{l}{0.5\textwidth}
    \vspace*{-0.2in}
    \includegraphics[scale=0.15]{tx_composer.png}
\end{wrapfigure}

\noindent
actions can be added to a transaction before it is applied.
A user can then apply this transaction and in the example above this would result in the state pane showing a single payment and in addition the contract in the Marlowe editor will have been reduced to \lstinline{Close}.

@@

An earlier paper~\cite{cryptoeprint:2016:1156} reviews smart contracts on blockchain. Here we look at a number of recent systems that bear direct comparison with Marlowe, rather than general purpose languages.
Nxt~\cite{Nxt}, is special-purpose in providing a ``fat'' high-level
API, containing built-in transaction types and transactions that support some 250 primitive operations; these can be ``scripted'' in a client (only) using a binding to the API, which is available, for instance, in JavaScript. In providing such specificity this bears comparison with our implementation of contracts from the ACTUS standard~\cite{actus}.

Our work is inspired by the original work of Peyton Jones and others~\cite{PeytonJones:2000} to describe financial contracts using a DSL embedded in Haskell. The Findel project~\cite{findel} examines financial contracts on the Ethereum platform, and is also based on~\cite{PeytonJones:2000}. The authors note that payments need to be bounded; this is made concrete in our
account by our notion of commitments. They take no account of commitments or timeouts as our approach does, and so are unable to guarantee some properties -- such as a finite lifetime -- built into Marlowe by design.

BitML~\cite{BitML} is a DSL for specifying Marlowe-like contracts that regulate transfers on the Bitcoin blockchain, and is implemented via a compiler that translates contracts into  Bitcoin transactions plus strategies. Participants execute a contract by appending these transactions on the Bitcoin blockchain, according to their strategies, which involve the exchange of bitstrings that guarantee to a very high probability the correctness of contract execution. Marlowe is directly implemented by an interpreter which could also be implemented on a covenant-based~\cite{covenants} extension of the Bitcoin blockchain.
